POPLMark goes Concurrent!\\Challenge problems for process calculi and behavioural types

### 2023-10-15 challenges/linearity.tex
first draft of complete proof of preservation of \scong

\end{proof}
\begin{lemma}[Preservation for $\equiv$]
  \label{le:presequiv}
  If \( \types{\Gamma;\Delta}{P} \) and \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{Q} \).
\end{lemma}
\begin{proof}
  By induction on the structure of a process context involved in the congruence. In the base case where the context is a hole, proceed by case analysis on the \TirName{Sc} rule applied:
  \begin{description}
  \item[\TirName {Par-Comm/Assoc}] By rearranging sub-derivations noting that
    order does not matter for linear contexts
  \item[\TirName{Par-Inact}]
    % right-to-left by strengthening~\ref{lemma:strenD}. Vice-versa,
    By picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
  \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear and applying weak and strenghening accordingly.
  \item[\TirName{Res-Inact:}] use weakening~\ref{lemma:weak} part 2
  \item[\TirName{Res}] Noting that order does not matter
  \end{description}
  For the step case, TBD
\end{proof}

>>>>

\end{proof}

\newcommand{\sconga}[2]{#1 \stackrel{a}{\equiv} #2}
To prove that congruence preserves typing we need to spell out in more detail the former. First, we denote with $\sconga\cdot\cdot$ the relation induced by the six axioms.

\begin{lemma}[Preservation for $\stackrel{a}{\equiv}$]
  \label{le:presequiva}
  If \( \sconga{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
\end{lemma}
\begin{proof}
  By case analysis on the \TirName{Sc} rule applied:
  \begin{description}
  \item[\TirName {Par-Comm/Assoc}] By rearranging sub-derivations noting that
    order does not matter for linear contexts
  \item[\TirName{Par-Inact}]
   Right-to-left by strengthening~\ref{lemma:strenD}. Vice-versa,
    by picking $T$ to be $\Tend$ and weakening~\ref{lemma:weak} part 2.
  \item[\TirName{Res-Par}] By case analysis on $x : T$ being linear or $\tend$ and applying weakening and strengthening accordingly.
  \item[\TirName{Res-Inact:}] Use weakening~\ref{lemma:weak} part 2
  \item[\TirName{Res}] Noting that order does not matter
  \end{description}
\end{proof}

Now, following Sangiorgi \& Walker define as the compatible equivalence relation induced by $\sconga\cdot\cdot$, namely add to reflexivity, symmetry  and transitivity the following:
 \begin{mathpar}
   \inferrule[Cong]{\sconga P Q}{\scong {C[P]}{C[Q]}}
 \end{mathpar}

\begin{lemma}[Preservation for $\equiv$]
  \label{le:presequiv}
    If \( \scong{P}{Q} \) then \( \types{\Gamma;\Delta}{P} \)  iff \( \types{\Gamma;\Delta}{Q} \).
\end{lemma}
\begin{proof}
  By induction on the structure of the derivation of \( \scong{P}{Q} \), with an inner induction of the structure of a process context.
  \begin{description}
  \item[\TirName {Refl}] Immediate
      \item[\TirName {Sym}] By IH.
      \item[\TirName {Trans}]  By two appeal to the IH
          \item[\TirName {Cong}] By induction on the structure of $C$. If the context is a hole, apply Lemma~\ref{le:presequiva}. In the step case, apply the IH: for example if $C$ has the form ${\Ppar C R}$, noting that ${\PBpar C R}[P]$ is equal to ${\Ppar {C[P]} R}$ we have \( \types{\Gamma;\Delta}{{\PBpar C R}[P]} \)  iff \( \types{\Gamma;\Delta}{{\PBpar C R}[Q]} \) by rule T-Par and the IH.
  \end{description}
\end{proof}
